---
title: "task01"
author: "Yirun Wang"
date: "16/08/2020"
output: html_document
---

```{r, message=FALSE}
library(tidyverse)
library(naniar)
library(data.table)
library(ggplot2)
library(ggmosaic)
library(readr)
```


## 0. Read in datasets
```{r}
transac <- fread("QVI_transaction_data.csv")
customer <- fread("QVI_purchase_behaviour.csv")

# head(transac)
```


## 1. EDA

### 1.1 Data Structures
```{r}
# transac -> date objects
transac$DATE <- as.Date(transac$DATE, origin = "1899-12-30")
```

### 1.2 Text Analysis: Are these products all chips?
```{r}
# explore the digits and special chars
prod_words <- data.table(unlist(strsplit(unique(transac[, PROD_NAME]), "")))
setnames(prod_words, 'words')


# remove digits and special chars
prods <- gsub("[[:punct:]]", "", transac$PROD_NAME)
prods <- gsub('[0-9]+', "", prods)

transac <- transac %>% 
  mutate(`prodname` = prods)

# sort by freq
chip_rank <- sort(table(prods), decreasing = TRUE)
```

### 1.3 Remove salsa
```{r}
transac <- transac %>% 
  mutate(`SALSA` = grepl("salsa", tolower(prodname)))

transac <- transac[which(transac$`SALSA` == FALSE),]

to_drop <- c("SALSA")
transac <- transac[, !(names(transac) %in% to_drop)]
```

### 1.4 Initial Summary: Check for nulls and outliers

No missing vals.

Outliers
- 200 packs of chips were purchased in a single transaction
- 650 total sales (occurred in the same transaction)

```{r}
## missingness (NB: uncomment next line to run the code - takes a while)
# vis_miss(transac, warn_large_data = FALSE)

# outliers
transac %>% 
  select_if(is.numeric) %>% 
  pivot_longer(c(PROD_QTY, TOT_SALES), names_to = "variable", values_to = "value") %>% 
  ggplot(aes(value), x = variable) +
  facet_wrap(~ variable, scales = "free") +
  geom_boxplot() +
  theme_bw(base_size = 18) +
  theme(strip.text.x = element_text(size = 8))

# max(transac$PROD_QTY)

# summary: means and medians
# transac %>% 
#   select_if(is.numeric) %>% 
#   pivot_longer(c(PROD_QTY, TOT_SALES), names_to = "variable", values_to = "value") %>% 
#   group_by(variable) %>%
#   summarise(mean = mean(value, na.rm = TRUE), median = median(value, na.rm = TRUE))
```

### 1.5.1 Filter the dataset: Pinpoint the transaction of interest

Two instances in which 200 packs of chips were purchased in one go.
- same customer (same loyalty #)
- same store
- same type of chips
- different dates (autumn and winter)

```{r}
transac %>% 
  filter(PROD_QTY == max(PROD_QTY))

# transac %>% 
#   filter(TOT_SALES == max(TOT_SALES))
```

### 1.5.2 Filter the dataset: Pinpoint the customer of interest

The customer has had no other recorded transactions. Chips may be purchased for commercial purposes.
Remove this cusomer as it does not yield insights into behaviour.

```{r}
transac %>% 
  filter(LYLTY_CARD_NBR == "226000")
```

### 1.6 Remove outliers

```{r}
transac <- transac[!transac$LYLTY_CARD_NBR == "226000", ]

# re-examine the dataset
# max(transac$PROD_QTY)
```

### 1.7 Count the number of transactions by date

Only 364 dates -> one day missing.

```{r}
transac_date <- transac %>% 
  group_by(DATE) %>% 
  summarise(freq = n())
```

### 1.8 Fill in the missing date
```{r}
# Create a sequence of unique dates from 1/7/18 to 30/6/19
unique_dates <- list(DATE = seq(as.Date('2018-07-01'), as.Date('2019-06-30'), by=1))
# class(unique_dates)

transac_yearly = merge(x=transac_date, y=unique_dates, by="DATE", all.y = TRUE)

# replace NA with 0
# transac_yearly <- transac_yearly %>% 
#   mutate_if(is.integer, ~replace(., is.na(.), 0))
```

### 1.9 Plot transactions over time
```{r}
# call theme_bw first ow previous settings changed

fig1 <- transac_yearly %>% 
  ggplot(aes(x = DATE, y = freq)) +
  geom_line() +
  labs(x = "Day", y = "Number of transactions", title = "Transactions over time") +
  scale_x_date(breaks = "1 month") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, vjust = 0.5))

fig1
```

### 2.0 December transactions

Increase -> ZERO on Xmas day due to the public holiday closure -> Decrease.

```{r}
fig2 <- transac_yearly[months(transac_yearly$DATE) %in% month.name[12],] %>% 
  ggplot(aes(x = DATE, y = freq)) +
  geom_line() +
  labs(x = "Day", y = "Number of transactions", title = "Transactions over time") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, vjust = 0.5))

fig2
```


## 2. Pack Size

### 2.1 Extract the digits
```{r}
# !!! parse_number

transac <- transac %>% 
  mutate(`PACK SIZE` = parse_number(PROD_NAME))

# Check outputs
c(max(transac$`PACK SIZE`), min(transac$`PACK SIZE`))
```

### 2.2 Plot pack size
```{r}
transac_pack <- transac %>% 
  group_by(`PACK SIZE`) %>% 
  summarise(freq = n())

# max(transac_pack$freq)

fig3 <- transac_pack %>% 
  ggplot(aes(x=`PACK SIZE`, y=freq)) +
  geom_bar(stat = "identity") +
  labs(x = "Pack Size", y = "Number of transactions", title = "Transactions by pack size") +
  theme_bw()

fig3
```

## 3. Brands

###





